### 2025-08-05 conteudo/background/codeduplication.tex
Review background

Code Duplication Detection is a field in computer science that has attracted
researchers' attention since 1988 \citep{firstman}. Code
duplication is considered a harmful artifact in software, as it negatively
affects tasks such as code readability and increases the likelihood of
introducing bugs \citep{harmone}. In most cases, code duplication tends to make
software more unstable than its non-duplicated counterpart
\citep{harmtwo}.

Over the years, research on Code Duplication Detection has branched into two
main areas: Code Clone Detection and Code Plagiarism Detection. The former
focuses on technical aspects, while the latter incorporates social dimensions
into the research field \citep{litreview}. In this work, we focus solely on the
Code Clone Detection branch due to the nature of our study.

\subsection{Types of Code Clone Duplication}

\label{subsec:types}

Classifying two code artifacts as duplicates raises the question of
what constitutes duplication. To address this, the literature categorizes code
duplication based on the extent of differences between code artifacts,
resulting in four types of code clone duplications \citep{litreview}. When
comparing two code artifacts, F and G, the four types of code duplication can
be described as follows \citep{litreview}:

\begin{itemize}
    \begin{item}
	\textbf{Type-1 (T1):} The differences between F and G are limited to
	    changes in comments, variable names, white spaces, and other
	    irrelevant elements. Figure \ref{fig:type1} shows a typical example
	    of Type-1 clone duplication \citep{litreview}.
    \end{item}
    \begin{item}
	\textbf{Type-2 (T2):} The differences between F and G include those
	    from Type-1, along with the addition and deletion of redundant
	    code. Figure \ref{fig:type2} shows a typical example of Type-2 code
	    clone duplication \citep{litreview}.
    \end{item}
    \begin{item}
	\textbf{Type-3 (T3):} The differences between F and G include those
	    from Type-2, as well as reordering code blocks and statements
	    within code blocks. Figure \ref{fig:type3} shows a typical example
	    of Type-3 code clone duplication \citep{litreview}.
    \end{item}
    \begin{item}
	\textbf{Type-4 (T4):} The differences between F and G include those
	    from Type-3, along with changes in data structures, the order of
	    operands/operators in expressions, or the replacement of code parts
	    with equivalent compositions. Figure \ref{fig:type4} shows a
	    typical example of Type-4 code clone duplication \citep{litreview}.
    \end{item}
\end{itemize}

\begin{figure}
\includegraphics[scale=0.38]{types/type1}
\caption{Example of Type-1 code clone duplication.}
The example is based on the work presented by CHEN et al. \citep{litreview}.
\label{fig:type1}
\end{figure}

\begin{figure}
\includegraphics[scale=0.38]{types/type2}
\caption{Example of Type-2 code clone duplication.}
The example is based on the work presented by CHEN et al. \citep{litreview}.
\label{fig:type2}
\end{figure}

\begin{figure}
\includegraphics[scale=0.38]{types/type3}
\caption{Example of Type-3 code clone duplication.}
The example is based on the work presented by CHEN et al. \citep{litreview}.
\label{fig:type3}
\end{figure}

\begin{figure}
\includegraphics[scale=0.38]{types/type4}
\caption{Example of Type-4 code clone duplication.}
The example is based on the work presented by CHEN et al. \citep{litreview}.
\label{fig:type4}
\end{figure}

@@

\begin{itemize}
    \item \textbf{Textual-based approaches:} According to Chen’s literature 
    review \citep{litreview}, this is the earliest methodology explored in 
    the field. It treats code as a text artifact and applies textual similarity
    techniques. An example is the work of Roy and Cordy \citep{textexample}, 
    which uses parsing and grammar techniques to detect code clones.

    \item \textbf{Token-based approaches:} This method tackles the problem 
    similarly to lexical analysis in the compilation process \citep{litreview}. 
    It involves identifying programming language tokens such as constants and 
    keywords, mapping variables to independent representations, etc. An example
    is the work of Toomey et al. \citep{tokenexample}, which detects code 
    clones through hashed token sequences.

    \item \textbf{Tree-based approaches:} While token-based approaches utilize
     lexical analysis, tree-based approaches rely on syntax trees, which store
     semantic information of code artifacts based on the programming language
     \citep{compiler}. An example of this approach is the work by Chilowicz 
     and Duris \citep{treeexample}.

    \item \textbf{Metric-based approaches:} This methodology uses extracted 
    metrics from code artifacts as embedding representations to measure 
    similarity \citep{litreview}, akin to the Word2Vec model by Mikolov et al. 
    \citep{wordtovec}. Metrics can include program size, number of variables, 
    memory access counts, etc. An example is the work by Kaur and Sharma \citep{metricexample}.

    \item \textbf{Graph-based approaches:} Similar to tree-based approaches, 
    this methodology depends on intermediate program representations, 
    specifically the program dependence graph (PDG), which stores information 
    on control, statements, and data dependencies \citep{prodg}. A recent 
    example is the work by Liu et al. \citep{tailor}, which represents one of 
    the state-of-the-art works in the field.
\end{itemize}

State-of-the-art code clone detection methods typically require substantial
computational time and memory, and many are language-specific, limiting their
applicability across different programming languages \citep{litreview}. Given
these constraints, we opted for a generalist textual-based approach using
Natural Language Processing (NLP) text similarity methods, which we will
introduce in Chapter \ref{cha:tool} and evaluate in this work.

@@

To evaluate and compare code clone detection methods, two widely adopted
datasets are available: OJ-Clone, collected from a pedagogical online judge
system \citep{ojclone}, and BigCloneBench, compiled from over 25,000 Java
software systems \citep{bigclonebench}. However, literature has expressed
concerns over the limited scope of OJ-Clone and the issues with BigCloneBench
identified by Krinke and Ragkhitwetsagul \citep{bigfail}. Despite these
limitations, BigCloneBench is commonly used due to the limited pedagogical
scope of OJ-Clone.

BigCloneBench further divides Type-3 duplications into three categories—Weak
Type-3 (WT3), Medium Type-3 (MT3), and Strong Type-3 (ST3)—where WT3 is closer
to Type-4 and ST3 is closer to Type-2 \citep{bigclonebench}. This results in
six possible outcomes for a code pair: Type-1, Type-2, Strong Type-3, Medium
Type-3, Weak Type-3/Type-4, and non-duplicate \citep{bigclonebench}.

@@

Using the counts of TP, FP, TN, and FN, standard metrics such as Precision,
Recall, and F-Score can be calculated \citep{recall}. In this work, we focus on
the Recall metric, which is commonly reported in literature for comparing code
clone detection approaches. Recall is defined as follows \citep{recall}:

>>>>

Code Duplication Detection is a field in computer science that has attracted
researchers' attention since 1988~\citep{firstman}. Code
duplication is considered a harmful artifact in software, as it negatively
affects tasks such as code readability and increases the likelihood of
introducing bugs~\citep{harmone}. In most cases, code duplication tends to make
software more unstable than its non-duplicated counterpart~\citep{harmtwo}.

Over the years, research on Code Duplication Detection has branched into two
main areas: Code Clone Detection and Code Plagiarism Detection. The former
focuses on technical aspects, while the latter incorporates social dimensions
into the research field~\citep{litreview}. In this work, we focus solely on the
Code Clone Detection branch due to the nature of our study.

\subsection{Diff command}

The \textit{diff} command is a command line in GNU/Linux that allows the user to compare 
two files line by line~\citep{diffcommand}. There are multiple forms in which the command 
results can be shown to the user, and Figure \ref{fig:diff} shows one of them. 
With the command results, it is easy to visualize and extract the number of lines that 
are equal and different between the two compared files.

\begin{figure}
\includegraphics[scale=0.2]{diff_command}
\caption[Example of \textit{diff} command executed]{
Example of \textit{diff} command executed
\footnote{
File 1 content is shown in red, and File 2 content is shown in green. \textit{-} 
represents a line that only exists in file 1, \textit{+} represents a line that only 
exists in file 2,\textit{!} represents a line that exists in both files but has modifications 
between them, and empty at the start of a line represents that the line exists in 
both files without any modifications.
}
}
\label{fig:diff}
\end{figure}

We use the \textit{diff} command as the base of one of the code duplication detection 
methods on the tool proposed in this work, although it is not explored as the text 
similarity method chosen for code duplication detection.

\subsection{Types of Code Clone Duplication}
\label{subsec:types}

Classifying two code artifacts as duplicates raises the question of
what constitutes duplication. To address this, the literature categorizes code
duplication based on the extent of differences between code artifacts,
resulting in four types of code clone duplications~\citep{litreview}. When
comparing two code artifacts, F and G, the four types of code duplication can
be described as follows~\citep{litreview}:

\begin{itemize}
    \begin{item}
	\textbf{Type-1 (T1):} The differences between F and G are limited to
	    changes in comments, variable names, white spaces, and other
	    irrelevant elements. Figure~\ref{fig:type1} shows a typical example
	    of Type-1 clone duplication~\citep{litreview}.
    \end{item}
    \begin{item}
	\textbf{Type-2 (T2):} The differences between F and G include those
	    from Type-1, along with the addition and deletion of redundant
	    code. Figure~\ref{fig:type2} shows a typical example of Type-2 code
	    clone duplication~\citep{litreview}.
    \end{item}
    \begin{item}
	\textbf{Type-3 (T3):} The differences between F and G include those
	    from Type-2, as well as reordering code blocks and statements
	    within code blocks. Figure~\ref{fig:type3} shows a typical example
	    of Type-3 code clone duplication~\citep{litreview}.
    \end{item}
    \begin{item}
	\textbf{Type-4 (T4):} The differences between F and G include those
	    from Type-3, along with changes in data structures, the order of
	    operands/operators in expressions, or the replacement of code parts
	    with equivalent compositions. Figure~\ref{fig:type4} shows a
	    typical example of Type-4 code clone duplication~\citep{litreview}.
    \end{item}
\end{itemize}

\begin{figure}[ht]
\includegraphics[scale=0.38]{types/type1}
\caption{Example of Type-1 code clone duplication~\citep{litreview}.}
\label{fig:type1}
\end{figure}

\begin{figure}[ht]
\includegraphics[scale=0.38]{types/type2}
\caption{Example of Type-2 code clone duplication~\citep{litreview}.}
\label{fig:type2}
\end{figure}

\begin{figure}[ht]
\includegraphics[scale=0.38]{types/type3}
\caption{Example of Type-3 code clone duplication~\citep{litreview}.}
\label{fig:type3}
\end{figure}

\begin{figure}[ht]
\includegraphics[scale=0.38]{types/type4}
\caption{Example of Type-4 code clone duplication~\citep{litreview}.}
\label{fig:type4}
\end{figure}

@@

\begin{itemize}
    \item \textbf{Textual-based approaches:} According to Chen’s literature 
    review \citep{litreview}, this is the earliest methodology explored in 
    the field. It treats code as a text artifact and applies textual similarity
    techniques. An example is the work of~\cite{textexample}, 
    which uses parsing and grammar techniques to detect code clones.

    \item \textbf{Token-based approaches:} This method tackles the problem 
    similarly to lexical analysis in the compilation process \citep{litreview}. 
    It involves identifying programming language tokens such as constants and 
    keywords, mapping variables to independent representations, etc. An example
    is the work of~\cite{tokenexample}, which detects code 
    clones through hashed token sequences.

    \item \textbf{Tree-based approaches:} While token-based approaches utilize
     lexical analysis, tree-based approaches rely on syntax trees, which store
     semantic information of code artifacts based on the programming
     language~\citep{compiler}. An example of this approach is the work of~\cite{treeexample}.

    \item \textbf{Metric-based approaches:} This methodology uses extracted 
    metrics from code artifacts as embedding representations to measure 
    similarity~\citep{litreview}, akin to the Word2Vec model by~\cite{wordtovec}.
    Metrics can include program size, number of variables, 
    memory access counts, etc. An example is the work of~\cite{metricexample}.

    \item \textbf{Graph-based approaches:} Similar to tree-based approaches, 
    this methodology depends on intermediate program representations, 
    specifically the program dependence graph (PDG), which stores information 
    on control, statements, and data dependencies~\citep{prodg}. A recent 
    example is the work of~\cite{tailor}, which represents one of 
    the state-of-the-art works in the field.
\end{itemize}

State-of-the-art code clone detection methods typically require substantial
computational time and memory, and many are language-specific, limiting their
applicability across different programming languages~\citep{litreview}. Given
these constraints, we opted for a generalist textual-based approach using
Natural Language Processing (NLP) text similarity methods, which we will
introduce in Chapter~\ref{cha:tool} and evaluate in this work.

@@

To evaluate and compare code clone detection methods, two widely adopted
datasets are available: OJ-Clone, collected from a pedagogical online judge
system~\citep{ojclone}, and BigCloneBench, compiled from over 25,000 Java
software systems~\citep{bigclonebench}. However, literature has expressed
concerns over the limited scope of OJ-Clone and the issues with BigCloneBench
identified by~\cite{bigfail}. Despite these
limitations, BigCloneBench is commonly used due to the limited pedagogical
scope of OJ-Clone.

BigCloneBench further divides Type-3 duplications into three categories: Weak
Type-3 (WT3), Medium Type-3 (MT3), and Strong Type-3 (ST3); where WT3 is closer
to Type-4 and ST3 is closer to Type-2~\citep{bigclonebench}. This results in
six possible outcomes for a code pair: Type-1, Type-2, Strong Type-3, Medium
Type-3, Weak Type-3/Type-4, and non-duplicate~\citep{bigclonebench}.

@@

Using the counts of TP, FP, TN, and FN, standard metrics such as Precision,
Recall, and F-Score can be calculated~\citep{recall}. In this work, we focus on
the Recall metric, which is commonly reported in literature for comparing code
clone detection approaches. Recall is defined as follows~\citep{recall}:

### 2025-08-05 conteudo/background/linux.tex
Review background

\section{The Linux Kernel}

The Linux kernel, a Unix-like operating system kernel, was published by Linus
Torvalds in 1992 as free software. It began gaining popularity in the late
1990s \citep{linuxbook}. Contrary to popular belief, Linux is not a
complete operating system, as it lacks components like filesystem utilities and
windowing systems. The so-called ``Linux distributions'' commonly
recognized by the public are GNU/Linux operating systems, which
include various essential applications alongside the Linux kernel,
such as system
\citep{gnuref}. Examples of Linux distributions are Ubuntu \citep{ubuntu} 
and Arch Linux \citep{archlinux}.

The Linux kernel is monolithic and organized into subsystems, such as
the process scheduler, memory management, device driver infrastructure,
networking, and filesystems \citep{melissa}. Each subsystem typically has a
dedicated maintainer or a team of maintainers responsible for overseeing its
development and managing contributions \citep{melissa}.

Development contributions to the Linux kernel are coordinated using the Git
version control system. These contributions are formatted as patches—text
documents that outline the differences between two source code versions.
These patches are then submitted and reviewed via the Linux mailing lists
\citep{melissa}.

Mailing lists are the primary means of communication within the Linux
community. Community members use them to submit and review patches,
discuss Linux-related topics, and debate potential new features. Given the
kernel's extensive modularity, each subsystem has its mailing list, allowing
discussions to be organized around specific areas of development
\citep{melissa}. All interactions on these mailing lists are documented and
archived in the Linux Kernel Lore \citep{linuxlore}, which serves as a
repository for kernel discussions and can be accessed by anyone interested.

One of the subsystems within the Linux kernel is the AMD Display driver, which
is responsible for implementing the drivers needed for AMD GPUs to function
correctly in a Linux environment. The maintainers of this subsystem have
reported a practical issue: a significant amount of duplicated code, which
complicates the maintenance of this subsystem. This problem serves as the
starting point for our research, with the AMD Display driver as our primary
object of study as we aim to understand and mitigate code duplication in the
Linux kernel. The AMD Display driver subsystem contains more than 391000 lines 
of code and more than 1089 code files
\footnote{ 
Data extracted directly from the Linux repository with the help of the cloc tool.
\citep{cloc}
}, 
making it a considerable component within the Linux kernel. Like other 
subsystems, discussions about the AMD Display driver are conducted on its 
designated mailing list, which can be accessed here: 
\url{https://lists.freedesktop.org/mailman/listinfo/amd-gfx}.

Another subsystem within the Linux kernel explored in this work is the
Industrial I/O subsystem (IIO). According to the Linux kernel documentation, 
''The main purpose of the Industrial I/O subsystem (IIO) is to provide support 
for devices that in some sense perform either analog-to-digital conversion (ADC) 
or digital-to-analog conversion (DAC) or both''\citep{iiodoc}. The IIO subsystem 
contains more than 281772 lines of code and more than 755 code files
\footnote{
Data extracted directly from the Linux repository with the help of the clol tool. \citep{cloc}
}.
We opted to include the IIO subsystem in our experiments to ensure our findings 
are not limited to one subsystem. Additionally, we have an established point of
contact within the IIO subsystem to facilitate this work.
Like other subsystems, discussions about the IIO subsystem are conducted on its 
designated mailing list, which can be found here: 
\url{https://subspace.kernel.org/vger.kernel.org.html}.

\subsection{Diff command}

The \textit{diff} command is a command line in Linux that allows the user to compare 
two files line by line \citep{diffcommand}. There are multiple forms in which the command 
results can be shown to the user, and Figure \ref{fig:diff} shows one of them. 
With the command results, it is easy to visualize and extract the number of lines that 
are equal and different between the two compared files.

\begin{figure}
\includegraphics[scale=0.25]{diff_command}
\caption[Example of \textit{diff} command executed]{
Example of \textit{diff} command executed
\footnote{
File 1's content is shown in red, and file 2's content is shown in green. \textit{-} 
represents a line that only exists in file 1, \textit{+} represents a line that only 
exists in file 2,\textit{!} represents a line that exists in both files but has modifications 
between them, and empty at the start of a line represents that the line exists in 
both files without any modifications.
}
}
\label{fig:diff}
\end{figure}

We use the \textit{diff} command as the base of one of the code duplication detection 
methods on the tool proposed in this work, although it is not explored as the text 
similarity method chosen for code duplication detection.

>>>>

\section{The Linux Kernel}
Linus Torvalds released the initial version of the Linux kernel, a Unix-like operating system kernel, in 1991. It was initially released under a license of Torvalds' creation, but the Linux kernel was later relicensed under GNU General Public License version 2 (GPLv2) in 1992.
%
It began gaining popularity in the late
1990s~\citep{linuxbook}. Contrary to popular belief, Linux is not a
complete operating system, as it lacks components like filesystem utilities and
windowing systems. The so-called ``Linux distributions'' commonly
recognized by the public are GNU/Linux operating systems, which
include various essential applications alongside the Linux kernel,
such as system~\citep{gnuref}. Examples of Linux distributions are Ubuntu 
and Arch Linux.

The Linux kernel is monolithic and organized into subsystems, such as
the process scheduler, memory management, device driver infrastructure,
networking, and filesystems~\citep{melissa}. Each subsystem typically has a
dedicated maintainer or a team of maintainers responsible for overseeing its
development and managing contributions~\citep{melissa}.

Development contributions to the Linux kernel are coordinated using the Git
version control system. These contributions are formatted as patches—text
documents that outline the differences between two source code versions.
These patches are then submitted and reviewed via the Linux mailing lists~\citep{melissa}.

Mailing lists are the primary means of communication within the Linux
community. Community members use them to submit and review patches,
discuss Linux-related topics, and debate potential new features. Given the
kernel extensive modularity, each subsystem has its mailing list, allowing
discussions to be organized around specific areas of development~\citep{melissa}. All interactions on these mailing lists are documented and
archived in the Linux Kernel Lore\footnote{\url{https://lore.kernel.org}}, which serves as a
repository for kernel discussions and can be accessed by anyone interested.

One of the subsystems within the Linux kernel is the AMD Display driver, which
is responsible for implementing the drivers needed for AMD GPUs to function
correctly in a Linux environment.
%
During the exploratory phase of this work, in a brainstorming conversation with the maintainers of this subsystem, they reported a practical issue: a significant amount of duplicated code, which complicates the maintenance of this subsystem.
%
This problem serves as the
starting point for our research, with the AMD Display driver as our primary
object of study as we aim to understand and mitigate code duplication in the
Linux kernel. The AMD Display driver subsystem contains more than 391,000 lines 
of code and more than 1,089 code files
\footnote{ 
Data extracted directly from the Linux repository with the help of the cloc tool~\citep{cloc}.
}, 
making it a considerable component within the Linux kernel. Like other 
subsystems, discussions about the AMD Display driver are conducted on its 
designated mailing list, which can be accessed here: 
\url{https://lists.freedesktop.org/mailman/listinfo/amd-gfx}.

Another subsystem within the Linux kernel explored in this work is the
Industrial I/O subsystem (IIO). According to the Linux kernel documentation, 
``The main purpose of the Industrial I/O subsystem (IIO) is to provide support 
for devices that in some sense perform either analog-to-digital conversion (ADC) 
or digital-to-analog conversion (DAC) or both''~\citep{iiodoc}. The IIO subsystem 
contains more than 281,772 lines of code and more than 755 code files
\footnote{
Data extracted directly from the Linux repository with the help of the clol tool~\citep{cloc}.
}.
We opted to include the IIO subsystem in our experiments to ensure our findings 
are not limited to one subsystem. Additionally, we have an established point of
contact within the IIO subsystem to facilitate this work.
Like other subsystems, discussions about the IIO subsystem are conducted on its 
designated mailing list, which can be found here: 
\url{https://subspace.kernel.org/vger.kernel.org.html}.

### 2025-08-05 conteudo/background/nlpsimilarity.tex
Review background

The standard methodology for text similarity detection is based on building
vector embeddings of texts and applying a distance function to these embeddings
to measure their similarity. The literature proposes multiple vector
embedding methods, such as bag-of-words, TF-IDF, LSI, and Word2Vec
\citep{gensimlivro}. Recent works related to vector embedding construction
utilize Large Language Models (LLMs), as exemplified by Ethayarajh's work,
which employs BERT, ELMo, and GPT \citep{llmsimilar}. The most popular distance
function used in the literature is cosine similarity, defined by the formula
\citep{cosineref}:

@@

It is well-known that LLMs have a high computational cost. For instance, an
analysis by Reimers and Gurevych found that finding the most similar pair in a
collection of 10,000 sentences required about 65 hours with BERT
\citep{bertsimilar}. Since we aim to compare pairs of similar magnitude in
our work, we decided not to use LLMs for our vector embedding method, even
though they may be suitable for smaller codebases.

@@

Gensim is an FLOSS Python library \citep{gensim}, which its creators
describe as "The fastest library for training of vector embeddings - Python or
otherwise. The core algorithms in Gensim use battle-hardened, highly optimized,
and parallelized C routines." \citep{gensimsite}. Due to Gensim's speed, we
chose this library for our vector embedding method.

>>>>

The standard methodology for text similarity detection is based on building
vector embeddings of texts and applying a distance function to these embeddings
to measure their similarity. The literature proposes multiple vector
embedding methods, such as bag-of-words, TF-IDF, LSI, and
Word2Vec~\citep{gensimlivro}. Recent works related to vector embedding construction
utilize Large Language Models (LLMs), as exemplified by Ethayarajh's work,
which employs BERT, ELMo, and GPT~\citep{llmsimilar}. The most popular distance
function used in the literature is cosine similarity, defined by the formula~\citep{cosineref}:

@@

It is well-known that LLMs have a high computational cost. For instance, an
analysis by Reimers and Gurevych found that finding the most similar pair in a
collection of 10,000 sentences required about 65 hours with
BERT~\citep{bertsimilar}. Since we aim to compare pairs of similar magnitude in
our work, we decided not to use LLMs for our vector embedding method, even
though they may be suitable for smaller codebases.

@@

Gensim is a FLOSS Python library \citep{gensim}, which its creators
describe as ``The fastest library for training of vector embeddings -- Python or
otherwise. The core algorithms in Gensim use battle-hardened, highly optimized,
and parallelized C routines''~\citep{gensimsite}. Due to Gensim speed, we
chose this library for our vector embedding method.
